---
title: "MUSA508_Ass2"
author: "Caesar"
date: "2023-09-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load package, warning = FALSE, message = FALSE} 
library(tidycensus)
library(tidyverse)
library(ggplot2)
library(sf)
library(kableExtra)

options(scipen=999)
options(tigris_class = "sf")

source("https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/functions.r")

palette5 <- c("#f0f9e8","#bae4bc","#7bccc4","#43a2ca","#0868ac")
```

## Load census data dictionaries


```{r load_variables, cache = TRUE}

acs_variable_list.2019 <- load_variables(2019, #year
                                         "acs5", #five year ACS estimates
                                         cache = TRUE)

acs_variable_list.2009 <- load_variables(2009, #year
                                         "acs5", #five year ACS estimates
                                         cache = TRUE)
```

### Use `get_acs()` to get 2019 ACS data

Notice this returns "long" data - let's examine it

```{r results='hide'}
tracts19 <-  
  get_acs(geography = "tract",
          variables = c("B25026_001E","B02001_002E",
                        "B15001_050E","B15001_009E",
                        "B19013_001E", "B25058_001E",
                        "B06012_002E"), 
          year=2019, state=11,
          geometry=TRUE) %>% 
  st_transform('ESRI:102728')
```

```{r}
glimpse(tracts19)
```

We create a new data frame consisting only of population

```{r}

totalPop19 <-
  tracts19 %>%
  filter(variable == "B25026_001")
```

Ways to examine the data

```{r}
nrow(totalPop19)

names(totalPop19)

head(totalPop19)

glimpse(totalPop19)
```

```{r}
A <- 
  ggplot() +
  geom_sf(data = totalPop19, aes(fill = estimate)) +
  theme(
    plot.title = element_text(size=22)
    )

B <- 
  ggplot() +
  geom_sf(data = totalPop19, aes(fill = q5(estimate))) +
  theme(plot.title = element_text(size=22)) 

C <-
  ggplot() +
  geom_sf(data = totalPop19, aes(fill = q5(estimate))) +
  scale_fill_manual(values = palette5,
                    labels = qBr(totalPop19, "estimate"),
                    name = "Total\nPopluation\n(Quintile Breaks)") +
  theme(plot.title = element_text(size=22))

D <- 
  ggplot() +
  geom_sf(data = totalPop19, aes(fill = q5(estimate))) +
  scale_fill_manual(values = palette5,
                    labels = qBr(totalPop19, "estimate"),
                    name = "Popluation\n(Quintile Breaks)") +
  labs(title = "Total Population", subtitle = "Washington DC; 2019") +
  mapTheme() + 
  theme(plot.title = element_text(size=22))
```

```{r}
# Let's "spread" the data into wide form

tracts19 <- 
  tracts19 %>%
  dplyr::select( -NAME, -moe) %>%
  spread(key = variable, value = estimate) %>%
  rename(TotalPop = B25026_001, 
         Whites = B02001_002,
         FemaleBachelors = B15001_050, 
         MaleBachelors = B15001_009,
         MedHHInc = B19013_001, 
         MedRent = B25058_001,
         TotalPoverty = B06012_002)


# Let's create new rate variables using mutate

tracts19 <- 
  tracts19 %>%
  mutate(pctWhite = ifelse(TotalPop > 0, Whites / TotalPop, 0),
         pctBachelors = ifelse(TotalPop > 0, ((FemaleBachelors + MaleBachelors) / TotalPop), 0),
         pctPoverty = ifelse(TotalPop > 0, TotalPoverty / TotalPop, 0),
         year = "2019") %>%
  dplyr::select(-Whites,-FemaleBachelors,-MaleBachelors,-TotalPoverty)
```

```{r results='hide'}
tracts09 <- 
  get_acs(geography = "tract", 
          variables = c("B25026_001E","B02001_002E",
                        "B15001_050E","B15001_009E",
                        "B19013_001E","B25058_001E",
                        "B06012_002E"), 
          year=2009, state=11, 
          geometry=TRUE, output="wide") %>%
  st_transform('ESRI:102728') %>%
  rename(TotalPop = B25026_001E, 
         Whites = B02001_002E,
         FemaleBachelors = B15001_050E, 
         MaleBachelors = B15001_009E,
         MedHHInc = B19013_001E, 
         MedRent = B25058_001E,
         TotalPoverty = B06012_002E) %>%
  dplyr::select(-NAME, -starts_with("B")) %>%
  mutate(pctWhite = ifelse(TotalPop > 0, Whites / TotalPop,0),
         pctBachelors = ifelse(TotalPop > 0, ((FemaleBachelors + MaleBachelors) / TotalPop),0),
         pctPoverty = ifelse(TotalPop > 0, TotalPoverty / TotalPop, 0),
         year = "2009") %>%
  dplyr::select(-Whites, -FemaleBachelors, -MaleBachelors, -TotalPoverty) 
```

```{r}
allTracts <- rbind(tracts19,tracts09)
```

```{r crimedata}
crime19 <- st_read("https://maps2.dcgis.dc.gov/dcgis/rest/services/FEEDS/MPD/MapServer/1/query?outFields=*&where=1%3D1&f=geojson") %>%
  dplyr::select(OFFENSE, CENSUS_TRACT, OBJECTID) %>%
  st_transform('ESRI:102728')  

crime09 <- st_read("https://maps2.dcgis.dc.gov/dcgis/rest/services/FEEDS/MPD/MapServer/33/query?outFields=*&where=1%3D1&f=geojson") %>%
  dplyr::select(OFFENSE, CENSUS_TRACT, OBJECTID) %>%
  st_transform('ESRI:102728') 
  
```
```{r crimewithin}
crime_counts_by_tract <- allTracts %>%
  st_intersection(crime19) %>%
  group_by(GEOID) %>%
  summarise(crime_counts=n()) %>%
  st_drop_geometry() %>%
  mutate(year="2019")

allTracts <- left_join(allTracts,crime_counts_by_tract,by=c("GEOID"="GEOID", "year"="year"))

```
```{r crimewithin}
crime_counts_by_tract <- allTracts %>%
  st_intersection(crime09) %>%
  group_by(GEOID) %>%
  summarise(crime_counts=n()) %>%
  st_drop_geometry() %>%
  mutate(year="2009")

allTracts <- left_join(allTracts,crime_counts_by_tract,by=c("GEOID"="GEOID", "year"="year"))

```

```


```{r }
dc_station <- st_read("https://maps2.dcgis.dc.gov/dcgis/rest/services/DCGIS_DATA/Transportation_Rail_Bus_WebMercator/MapServer/52/query?where=1%3D1&outFields=*&outSR=4326&f=json") %>%
  dplyr::select(NAME,LINE) %>%
  st_transform('ESRI:102728')  

dc_line <- st_read('https://maps2.dcgis.dc.gov/dcgis/rest/services/DCGIS_DATA/Transportation_Rail_Bus_WebMercator/MapServer/106/query?outFields=*&where=1%3D1&f=geojson') %>%
  dplyr::select(NAME,GIS_ID)%>%
  st_transform('ESRI:102728')
```

```{r}
ggplot() +
  geom_sf(data = st_union(tracts19)) +
  geom_sf(data = dc_station,
          aes(color = LINE),
          show.legend = 'Point',size=2) +
  geom_sf(data = dc_line,
          aes(color = GIS_ID),
          show.legend = 'Line',size=1)
  labs(title = 'Station stops',
       subtitle = 'Washington DC',
       caption = 'Figure 2.2')
```

```{r }
station_buffer<- rbind(
  st_buffer(dc_station,2640) %>%
    mutate(Legend = 'buffer')%>%
    dplyr::select(Legend),
  st_union(st_buffer(dc_station,2640))%>%
    st_sf()%>%
    mutate(Legend = 'Unioned Buffer')
)
```

```{r}
ggplot() +
  geom_sf(data = station_buffer) +
  geom_sf(data = dc_station,
          show.legend = 'Point',size = 2) +
  facet_wrap(~Legend) + 
  labs(caption = 'Figure 2.3')+
  mapTheme()
```

```{r}
buffer <- filter(station_buffer, Legend=='Unioned Buffer')
```

```{r}
clip <- st_intersection(buffer,tracts19) %>%
  dplyr::select(TotalPop)%>%
  mutate(inter_type = 'Clip')
```

```{r}
selection <- tracts19[buffer,]%>%
  select(TotalPop)%>%
  mutate(inter_type = 'Spatial Selection')
```

```{r}
select_centroid <- st_centroid(tracts19)[buffer,] %>%
  st_drop_geometry() %>%
  left_join(., dplyr::select(tracts19, GEOID), by = "GEOID") %>%
  st_sf() %>%
  dplyr::select(TotalPop) %>%
  mutate(inter_type = "Centroids")
```

```{r}
intersections <- rbind(clip, selection, select_centroid)

ggplot() +
  geom_sf(data=intersections, aes(fill = TotalPop)) +
  geom_sf(data=dc_station, show.legend = "point") +
  scale_fill_viridis_c() +
  facet_wrap(~inter_type) + 
  mapTheme()
```

```{r}
allTracts.group <- 
  rbind(
    st_centroid(allTracts)[buffer,] %>%
      st_drop_geometry() %>%
      left_join(allTracts) %>%
      st_sf() %>%
      mutate(TOD = "TOD"),
    st_centroid(allTracts)[buffer, op = st_disjoint] %>%
      st_drop_geometry() %>%
      left_join(allTracts) %>%
      st_sf() %>%
      mutate(TOD = "Non-TOD")) %>%
  mutate(MedRent.inf = ifelse(year == "2009", MedRent * 1.19, MedRent)) 
```